---
title: Thesis Title
subtitle: Subtitle
author: 
  - name: Bart Kudrzycki
    affiliation: Development Economics Group, ETH Zurich, Switzerland, <bartlomiej.kudrzycki@nadel.ethz.ch>
date: \today
thanks: Thanks to Dario and Elina for the Markdown template.
supervisors: Isabel Gunther
acknowledgments: I would like to thank...
bigabstract: Big abstract
keywords: [Informal labor markets, Dual training, Apprenticeship]
jel: [I26]
bibliography: bibliography.bib
csl: 'apa-no-ampersand.csl'
output: 
  bookdown::pdf_document2:
    latex_engine: xelatex
    template: template.tex
    extra_dependencies: ["flafter", "placeins", "changepage", "multirow", "array", "adjustbox", "longtable", "pifont", "fancyhdr", "pdflscape"]
---

\doublespacing

\chapter{Introduction}

\clearpage

\chapter{Lost in Transition}
\begin{chapabstract}
This paper uses a novel longitudinal dataset of 752 youth living in Cotonou, Benin, collected over 2 years, to analyze the school-to-work transition in a highly informal, urban environment. We conduct five waves of in-person and mobile phone surveys over three years with youth aged 20-29. We find evidence of queuing for wage jobs: relatively well-educated youth endure long periods of inactivity waiting for wage employment. Even wage jobs are informal and low-paying, and almost none fulfill the ILO formality criteria. Quantifying the transition, we find that the average youth in our sample needs between 5 and 18 months to find their first employment after graduation and between 23 and 27 months to find steady employment. Transition propensities show that casual, but full-time work following a spell of inactivity is the most common path taken; it is more likely to occur directly after graduation for men than from women. Finally, we find that youth who have transitioned to the labor market experience more life satisfaction, though no effect is detected for particular types of transitions.
\end{chapabstract}

\clearpage

```{r, child='chapters/1_survey/markdown/survey.Rmd'}
```

\newsection

\chapter{Youth Labor Index for Low Income Countries}
_co-authored with Erwin Lefoll & Isabel Günther_

\begin{chapabstract}
This paper introduces the Youth Labor Index for Lower-Income Countries (YLILI), a composite index of 10 labor market indicators specifically tailored to developing countries. The indicators are chosen based on their relevance for economies with high levels of informal work and because they can be computed for ages 15-24 and disaggregated by gender. The index rankings suggest that poor education outcomes and high working poverty rates are the main drivers of weak youth labor market outcomes, especially in sub-Saharan Africa. Labor market scores are higher for males, a result driven by the large shares of young women not in education or the workforce. The fertility rate is a strong predictor of index scores, suggesting that countries with fastest growing youth populations are also struggling to provide sufficient opportunities for decent youth employment. A webtool allows users to explore the YLILI and identify policy priorities.
\end{chapabstract}

\clearpage

```{r, child='chapters/2_index/markdown/index.Rmd'}
```

\newsection

\chapter{Costs and Benefits of the Dual System in the Context of Traditional Apprenticeship in Bénin}
_co-authored with Sylvain Kpenavoun, Esaïe Gandonou, Guy Nouatin & Rubain Bankole _
\begin{chapabstract}
Traditional apprenticeships are an important source of skills for early school leavers in developing countries. Usually private arrangements between parents and informal firms, apprenticeship quality and costs are not subject to oversight or regulation, and the various transactions taking place between apprentice and firm have not been carefully documented to date. This paper sheds light on the institution of traditional apprenticeship through a study of a national apprenticeship reform called the CQP, which combines in-firm training with a weekly classroom component. Using two waves of surveys with firm owners and apprentices, we analyze the human capital gains and material costs and benefits associated with training. We find that the program did not generate improvements in apprentice skill or in firm profitability, possibly due to a strong selection of older, overqualified candidates into the program. We do find that traditional apprenticeship on the whole improves apprentices' sector-specific knowledge and experience. Fees paid by apprentices are found to be commensurate with the allowances they receive from the trainers, contradicting earlier work which suggest that firms use fees for financing. Estimates suggest that apprentice productivity must be substantial for firms to break even on training. The high proportion of apprentices as a fraction of the total workforce in firms, combined with the tendency for apprentices to leave their training workshop after graduation, suggests that apprentice productive contributions to the training firms are indeed substantial.
\end{chapabstract}

\clearpage

```{r, child='chapters/3_cqp/markdown/cqp.Rmd'}
```

\chapter{Conclusion}