---
title: "Bridging the Gap: School-to-Work Transitions and Youth Labor Market Dynamics in Informal Economies"
dissno: 29728
author: 
  - name: Bartlomiej Pawel Kudrzycki
thanks: Thanks to Dario and Elina for the Markdown template.
keywords: [Informal labor markets, Dual training, Apprenticeship]
jel: [I26]
classoption: a4paper, twoside, 12pt
geometry: inner=3cm, outer=3cm
bibliography: bibliography.bib
biblatexoptions: [refsegment=chapter]
output: 
  bookdown::pdf_book:
    latex_engine: xelatex
    citation_package: biblatex
    keep_tex: yes
    link-citations: yes
    template: template.tex
    extra_dependencies: ["flafter", "placeins", "changepage", "multirow", "array", "adjustbox", "longtable", "pifont", "fancyhdr", "pdflscape", "fancyhdr", "threeparttable"]

---

```{r diss_setup, include=FALSE}
packages <- c("pandoc", "bookdown")

options(repos = c(CRAN = "http://cran.rstudio.com"))

# install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages], silent = TRUE)
}

# load packages
invisible(lapply(packages, library, character.only = TRUE))

rm(installed_packages)

```

\onehalfspacing
\raggedbottom

\chapter{Introduction}
```{r, child='chapters/0_intro/markdown/introduction.Rmd'}
```

\newsection

\chapter{Youth Labor Index for Low Income Countries}
_co-authored with Erwin Lefoll & Isabel Günther_

\begin{chapabstract}
This paper introduces the Youth Labor Index for Lower-Income Countries (YLILI), a composite index of 10 labor market indicators specifically tailored to developing countries. The indicators are chosen based on their relevance for economies with high levels of informal work and because they can be computed for ages 15-24 and disaggregated by gender. The index rankings suggest that poor education outcomes and high working poverty rates are the main drivers of weak youth labor market outcomes, especially in sub-Saharan Africa. Labor market scores are higher for males, a result driven by the large shares of young women not in education or the workforce. The fertility rate is a strong predictor of index scores, suggesting that countries with fastest growing youth populations are also struggling to provide sufficient opportunities for decent youth employment. A webtool allows users to explore the YLILI and identify policy priorities.
\end{chapabstract}

\newpage

```{r, child='chapters/1_index/markdown/index.Rmd'}
```

\newsection

\chapter[Lost in Transition]{Lost in Transition: School-to-Work Transition Mapping in Urban Bénin}

\begin{chapabstract}
This paper uses a novel, longitudinal dataset of 752 youth aged 20-29 in Cotonou, Benin, to map school-to-work transitions (SWTs) in a highly informal, urban economy. Five waves of in-person and mobile phone surveys conducted over three years provide rich employment and activity histories for youth. First, the activity histories are used to estimate the duration of the transition from school to employment. Second, transition matrices constructed from the event histories are used identify differences in rates of transition between activity states by gender and age group. Third, Optimal Matching Analysis (OMA) treats the activity histories as sequences, allowing us to group and analyze clusters of similar school-to-work trajectories. Fourth, we develop a detailed taxonomy of employment types (such as casual work or independent self-employment) and analyze the propensity for youth to engage in or transition into these types of work. We document limited permeation between activity states, a low probability of exiting the SWT among young women, and a late school-leaving age among all youth. We interpret the latter as indicative of strong sorting in the early stages of the SWT in informal labor markets.

\end{chapabstract}

\newpage

```{r, child='chapters/2_survey/markdown/survey.Rmd'}
```

\newsection

\chapter[Costs \& Benefits of Dual Apprenticeship in Bénin]{Costs and Benefits of the Dual System in the Context of Traditional Apprenticeship in Bénin}
_co-authored with Sylvain Kpenavoun Chogou, Esaïe Gandonou, Guy Nouatin & Rubain Bankole_

\begin{chapabstract}
Traditional apprenticeships are an important source of skills for early school leavers in developing countries. This paper sheds light on the institution of traditional apprenticeship through a study of a national apprenticeship reform called the CQP, which combines in-firm training with a weekly classroom component. Using two waves of surveys with firm owners and apprentices, we analyze the human capital gains and material costs and benefits associated with training. We find that the program did not generate improvements in apprentice skill or in firm profitability, possibly due to a strong selection of older, overqualified candidates into the program. We do find that traditional apprenticeship on the whole improves apprentices' sector-specific knowledge and experience. Fees paid by apprentices are found to be commensurate with the allowances they receive from the trainers, contradicting earlier work which suggest that firms use fees for financing. Estimates suggest that apprentice productivity must be substantial for firms to break even on training. The high proportion of apprentices as a fraction of the total workforce in firms, combined with the tendency for apprentices to leave their training workshop after graduation, suggests that apprentice productive contributions to the training firms are indeed substantial.
\end{chapabstract}

\newpage

```{r, child='chapters/3_cqp/markdown/cqp.Rmd'}
```
