I would like to thank Isabel Günther for her support throughout this dissertation. In the uncertain landscape of academic life, there are few guides as steady and generous, and I am deeply grateful for her wisdom, encouragement, and unflagging patience.

Individuals who have commented and contributed to various portions of this dissertation include my co-authors, Erwin Lefoll, Sylvain Kpenavoun Chogou, Esaïe Gandonou, Guy Nouatin and Rubain Bankole, my co-examiners Paola Monserrat Bordón Tapia and Ursula Renold, my TVET4INCOME co-conspirators Johanna Kemper, Jutta Buergi, Durga Prasad Baral, Miski Peralta, and Silvia Camacho Calvo, DEC researchers Kenneth Harttgen, Laura Metzger, Chris Humphrey, Patrick Premand, and Rahul Lahoti, my fellow doctoral students, Samuel Tetteh-Bah, Selina Bezzola, Yael Borofsky, Antoinette van der Merwe, Kathrin Durizzo, Dario Meili, and Frank Odhiambo, and seminar participants at the University of Zürich, the ETH, the University of Konstanz, the African School of Economics, the University of Abomey-Calavi, and multiple Zoom conferences. Your input and criticisms were invaluable in shaping the final work.

I would also like to extend a heartfelt thank you to all my collaborators in Bénin. Esaïe Gandonou and Guy Nouatin hosted me at the University of Abomey-Calavi, introduced me to key TVET policy-makers and ensured my safety and comfort during all three of my stays. Sylvain Kpenavoun Chogou was an enumerator trainer \textit{extraordinaire}, without whose efforts and sharp eye for detail I would have been left stranded in the field. The intrepid enumerators Ambroise, Appolinaire, Armel, Crésus, Emmanuel, Fiacre, Floriane, Nadège, Olayidé, and Serge collected the data which constitutes the bedrock of this work - I am deeply grateful for the quality of your surveying and the tirelessness of your investigative efforts. Finally, I am deeply indebted to Rubain, the best research collaborator, host, fellow traveler, and friend that I could ask for. Without your reliability and quick thinking, my field research would have faltered at inception. Without your endless optimism, infectious humor, and camaraderie, the PhD experience would have been incomparably less joyful. \textit{A wà nŭ kaka.}

A resounding thank you goes to all the wonderful people of NADEL: to Laura, my longest-running officemate, for the many shared laughs, cups of coffee, and words of encouragement. To Adina, who together with Laura and Isabel invited me to join the NADEL team. To my fellow PhDs Samuel, Selina, Yael, Antoinette, Kathrin, Dario, Erwin and Frank: your collective intelligence, curiosity, and enthusiasm for research have inspired and motivated me throughout the journey. To Hervé and Ilari, for never shying away from big questions, controversial ideas, and incommensurable theories. To Rebecca, Joschka, Johannes, and Nicolas, for sharing your considerable knowledge and experience. To Tabea, for making it possible for me to focus my time and energy on research (and reminding me to pay my library fees). To Kimon, for pushing me to my limits physically when I was already at them mentally. And to Marie-Laure, Fritz, Leonie, Elisabeth, Felicitas, Shruti, Evelyn, Bente, Mathias, Elina, and everyone else whose path I crossed over an immensely valuable 7 years.

Many others have supported me in various ways during the writing of this dissertation. My teachers at the ETH and the University of Zürich have stimulated my thinking and given me a taste of the breadth and depth of human understanding. My family and friends have exhibited remarkable patience during the challenges that are part and parcel of the dissertation process. My parents Kasia and Michał have been unwavering in their support from afar: you have been a constant source of encouragement and inspiration, and you taught me how to ask sensible questions long before I ventured out into the world. To your efforts and sacrifices I owe everything I that have found in it. My sister Alexandra has been a kindred spirit on this side of the pond: thank you for all the letters, visits, and calls.

Finally, my \textit{Lebensabschnittsgefährtin} Rowena: even gon Voethe couldn't find the words to thank you for your boundless love and support. This feat is as much yours as it is mine; I humbly dedicate it to your beautiful soul and persevering spirit.

\vspace{3cm}

\noindent This research was supported by Swiss Agency for Development and Cooperation and the Swiss National Science Foundation under the auspices of the r4d Research Program, as well as the Development Economics Group at the ETH.